---
title: "Data-Wrangle.rmd"
author: "Sanil Partha"
date: "10/17/2023"
output: html_document
---
```{r}
library(lubridate)
library(tidyverse)
library(readxl)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Creating a list of file names(which include the date of each file's data)
```{r}
time_series_list <- list.files('../Data/csse_covid_19_daily_reports_us', full.names = TRUE)

```
Reading in the first 18 columns (gets rid of 3 extra columns that data was not collected for in 2023) of each file in the list,

adding the date from the file name to a new "Date" column in each,

and adding a "Converted Date" column in the m-d-y format
```{r}
df <- lapply(time_series_list, function(file){
  table <- read.csv(file)[,1:18]
  table$Date <- substring(file, 40,49) # different character positions accounting for ../ extension in markdown environment
  table$Converted_Date <- as_date(table$Date, format = "%m-%d-%Y")

  table
})

head(df, n = 5)
```
Creates a new dataframe with all days' data stacked on top of each other using row bind and

Filters out territories and cruises that are not US states, and have no accompanying policy data

```{r}
binded_df <- do.call(rbind, df)
head(binded_df)
```

```{r}
binded_df <- filter(binded_df, !(binded_df$Province_State %in% c("American Samoa",
                                                             "Diamond Princess",
                                                             "Grand Princess",
                                                             "Guam",
                                                             "Northern Mariana Islands",
                                                             "Puerto Rico",
                                                             "Virgin Islands",
                                                             "Recovered")))

```


sorting and saving

```{r}
#sort from least recent to most recent 
sorted_df <- binded_df %>% arrange(ymd(binded_df$Converted_Date))

head(sorted_df)

#save cleaned data
write.csv(sorted_df, file = "../Data/csse_covid_19_daily_reports_us_CLEAN.csv")
```


Adding stringency index to new sorted dataframe, naming the new dataframe tidyData_v1

Each version of the tidyData table adds an index until all 4 are added in tidyData_v4

```{r}

#Joining in time series data - Stringency Index


StringencyIndex <- read_excel('../Data/OxCGRT_timeseries_us_subnational_indices.xlsx', 
                              sheet = "StringencyIndex")

StringencyIndex_clean <- StringencyIndex %>%
  filter(!is.na(RegionName)) %>%
  pivot_longer(cols= c(colnames(StringencyIndex[8:1162])),
                    names_to='Date',
                    values_to='StringencyIndex')

  
StringencyIndex_clean$Converted_Date <- as_date(StringencyIndex_clean$Date, format = "%d%b%Y")

StringencyIndex_clean$RegionName[StringencyIndex_clean$RegionName == "Washington DC"] <- "District of Columbia"

StringencyIndex_clean$Date <- NULL

tidyData_v1 <- left_join(sorted_df, 
                       StringencyIndex_clean, 
                       by = c("Province_State" = "RegionName", "Converted_Date"))



head(tidyData_v1)
```

Adding containment health index to v1, naming the new dataframe tidyData_v2:

```{r}
#Joining in time series data - ContainmentHealthIndex
ContainmentHealthIndex <- read_excel('../Data/OxCGRT_timeseries_us_subnational_indices.xlsx', 
                              sheet = "ContainmentHealthIndex")

ContainmentHealthIndex_clean <- ContainmentHealthIndex %>%
  filter(!is.na(RegionName)) %>%
  pivot_longer(cols= c(colnames(ContainmentHealthIndex[8:1162])),
               names_to='Date',
               values_to='ContainmentHealthIndex')

ContainmentHealthIndex_clean <- ContainmentHealthIndex_clean[,c(4,8,9)]

ContainmentHealthIndex_clean$Converted_Date <- as_date(ContainmentHealthIndex_clean$Date, format = "%d%b%Y")

ContainmentHealthIndex_clean$RegionName[ContainmentHealthIndex_clean$RegionName == "Washington DC"] <- "District of Columbia"

ContainmentHealthIndex_clean$Date <- NULL

tidyData_v2 <- left_join(tidyData_v1, 
                      ContainmentHealthIndex_clean, 
                      by = c("Province_State" = "RegionName", "Converted_Date"))


head(tidyData_v2)
```

Adding government response index to new sorted v2, naming the new dataframe tidyData_v3:

```{r}
#Joining in time series data - GovernmentResponseIndex
GovernmentResponseIndex <- read_excel('../Data/OxCGRT_timeseries_us_subnational_indices.xlsx', 
                                     sheet = "GovernmentResponseIndex")

GovernmentResponseIndex_clean <- GovernmentResponseIndex %>%
  filter(!is.na(RegionName)) %>%
  pivot_longer(cols= c(colnames(GovernmentResponseIndex[8:1162])),
               names_to='Date',
               values_to='GovernmentResponseIndex')

GovernmentResponseIndex_clean <- GovernmentResponseIndex_clean[,c(4,8,9)]

GovernmentResponseIndex_clean$Converted_Date <- as_date(GovernmentResponseIndex_clean$Date, format = "%d%b%Y")

GovernmentResponseIndex_clean$RegionName[GovernmentResponseIndex_clean$RegionName == "Washington DC"] <- "District of Columbia"

GovernmentResponseIndex_clean$Date <- NULL

tidyData_v3 <- left_join(tidyData_v2, 
                         GovernmentResponseIndex_clean, 
                         by = c("Province_State" = "RegionName", "Converted_Date"))


head(tidyData_v3)

```

Adding economic support index to new sorted v3, naming the new dataframe tidyData_v4:

```{r}
#Joining in time series data - EconomicSupportIndex
EconomicSupportIndex <- read_excel('../Data/OxCGRT_timeseries_us_subnational_indices.xlsx', 
                                      sheet = "EconomicSupportIndex")

EconomicSupportIndex_clean <- EconomicSupportIndex %>%
  filter(!is.na(RegionName)) %>%
  pivot_longer(cols= c(colnames(EconomicSupportIndex[8:1162])),
               names_to='Date',
               values_to='EconomicSupportIndex')

EconomicSupportIndex_clean <- EconomicSupportIndex_clean[,c(4,8,9)]

EconomicSupportIndex_clean$Converted_Date <- as_date(EconomicSupportIndex_clean$Date, format = "%d%b%Y")

EconomicSupportIndex_clean$RegionName[EconomicSupportIndex_clean$RegionName == "Washington DC"] <- "District of Columbia"

EconomicSupportIndex_clean$Date <- NULL

tidyData_v4 <- left_join(tidyData_v3, 
                         EconomicSupportIndex_clean, 
                         by = c("Province_State" = "RegionName", "Converted_Date"))


head(tidyData_v4)
```

Creating the daily deaths column

```{r}
dailyDeathsCount <- function(Deaths){
  d <- sapply(2:length(Deaths), function(i){
    Deaths[i] - Deaths[i-1]
  })
  
  c(Deaths[1],d)
  
}

tidyData_v5 <- tidyData_v4 %>%
  group_by(Province_State) %>%
  mutate(dailyDeaths = dailyDeathsCount(Deaths))

head(tidyData_v5)
```



saving tidy table

```{r}
#save cleaned data
write.csv(tidyData_v5, file = "../Data/covid19_state_policy_tidydata.csv")

```


